---
title: "Bayesian Model Averaging with the Laplace Approximation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Model Averaging with the Laplace Approximation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

BayesianId is a regression package designed to straightforwardly tell you the probabilities of hypotheses given data. In other words, it gives you the posterior probabilities of hypotheses about model parameters. These hypothesis probabilities are what researchers really want instead of p-values (in their "Bayesian id", Gigerenzer, 2004).
  The basic workflow for BayesianId is as follows
  1) Fit a model with all the predictors (independent variables) that you think might be important. This is the full model.
  2) Use plots of data against model predictions to assess the full model (distributional assumptions etc., NOT the set of predictors) and revise it if needed. For example, you might change from using basic linear regression to using log-linear regression.
  3) Use the "bma" function with the (possibly revised) full model to perform Bayesian model averaging. First, the code fits models with all possible subsets of predictors. Then it computes how probable each model is given the data, which is based on how well each model predicted the data. You can then make inferences based on the whole set of models - weighted by their probabilities - instead of just one "best" model.
  4) Analyze the "bma" object to make inferences about model parameters, which is what people typically care about.
  To better understand how BayesianId uses Bayesian model averaging for interpretable inference, please see the relevant vignette.
  BayesianId comes with two functions (breg_laplace and breg_mcmc) for estimating commonly used regression models (linear regression, logistic regression, ordinal logistic regression etc.). These run precompiled Stan code to estimate the models using either the Laplace approximation (which is much faster but potentially less accurate) or Markov chain Monte Carlo simulation (which is much slower but potentially more accurate). There is a vignette explaining how both functions (breg_laplace and breg_mcmc) are used. In most cases, I recommend using these built in functions for Bayesian model averaging.
  However if the model you want to estimate is not supported yet by breg_laplace/breg_mcmc, you do Bayesian model averaging with most models that have been estimated using maximum likelihood estimation (e.g. from the lm or glm functions). This uses the BIC approximation to compute model probabilities for Bayesian model averaging. Plugging in maximum likelihood estimates like this gives you "quick and dirty" approximate Bayesian inference that is still better than using p-values.

```{r setup}
library(BayesianId)
```

```{r load_data}
data("penguins")
head(penguins)
```

```{r full_model}
mass_breg_full = breg_laplace(#body_mass ~ island*sex + flipper_len + bill_len,
                              body_mass ~ sex,
                              data = na.omit(penguins),
                              prior_scale = 0.707)
```

```{r check_full_model}
plot(mass_breg_full, xvar = "bill_len")
```

```{r}
plot(mass_breg_full, group = "sex")
```


```{r create_bma_object}
tic = Sys.time()
mass_breg_bma = bma(mass_breg_full)
toc = Sys.time()
print(toc - tic)
```

```{r print}
print(mass_breg_bma)
```

```{r, contrast}
contrast(mass_breg_bma, factors = c("island", "sex"), ref = "Biscoe male")
```

```{r term_probs}
summary(mass_breg_bma, type = "term_probs")
```

```{r term_odds}
summary(mass_breg_bma, type = "term_odds")
```

```{r dir_probs}
summary(mass_breg_bma, type = "dir_probs")
```

```{r dir_odds}
summary(mass_breg_bma, type = "dir_odds")
```

```{r model_probs}
summary(mass_breg_bma, type = "model_probs")
```

```{r model_odds}
summary(mass_breg_bma, type = "model_odds")
```

```{r est}
summary(mass_breg_bma, type = "est")
```
